Trends and Challenges in Toronto’s Homeless Population (2018-2024)

### 2024-09-26 paper/paper.qmd
add a reference

---
title: "Trends and Challenges in Toronto’s Homeless Population (2018-2024)"
subtitle: "A Focus on Demographics and Chronic Homelessness"
author: 
  - Jianing Li
thanks: "Code and data are available at: https://github.com/JianingLi1225/Toronto_Homeless"
date: today
date-format: long
abstract: "This paper explores Toronto’s homeless population from 2018 to 2024, focusing on age, gender, and chronic homelessness trends. The data reveal a decline during the pandemic, followed by a rebound, with males and individuals aged 35-44 making up the majority. The number of chronically homeless also increased significantly. These findings highlight the need for targeted interventions to address chronic homelessness and improve housing policies."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#### Workspace setup ####
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
```

@@

In this essay, @sec-data-overview explains the background and necessity of using this dataset, followed by an explanation of the variables mentioned and the methods used. @sec-data-results shows the changes in the number of homeless individuals, along with the proportions and trends over time for each age group, gender, and chronic homelessness. Graphs and related explanations are provided. @sec-discussion further analyses the information presented in the graphs.

@@

All processes are performed through the `R` language [@citeR]. Data is simulated, tested, and cleaned using the `tidyverse` [@tidyverse] packages, which also facilitate downloading data through `opendatatoronto` [@citeopendatatorontoR]. Unneeded variables are removed and renamed using `dplyr` [@dplyr] and `tidyr` [@tidyr], while graphs are plotted with `ggplot2` [@ggplot2].

@@

@fig-total-population illustrates the fluctuation in Toronto's homeless population from 2018 to 2024. The chart shows cyclical rises and falls, with more pronounced declines around 2020 and 2021, followed by a steady increase.The population reaches its highest point in 2024. Throughout this period, the homeless population varies between approximately 7,500 and 11,000 individuals.

@@

@fig-chronic-population displays the changes in the number and proportion of chronic and non-chronic homeless people in Toronto between 2018 and 2024. The number of chronically homeless has been growing over this time, with over 30,000 people in 2018, and by 2023, the number has gone to about 60,000, nearly doubling. the slight drop in numbers in 2024 is due to the fact that the statistics were last updated on September 17, 2024, and it is projected that by the end of the year, the number will have exceeded the 2023 one. Similarly, the percentage of chronically homeless is growing every year, from 29.7% in 2019 to 63.7% in 2024. Note that in 2019 to 2020 and 2023 to 2024, the rate grows faster, by about 10 percent.

>>>>

---
title: "Trends and Challenges in Toronto’s Homeless Population (2018-2024)"
subtitle: "A Focus on Demographics and Chronic Homelessness"
author: 
  - Jianing Li
thanks: "Code and data are available at: https://github.com/JianingLi1225/Toronto_Homeless"
date: today
date-format: long
abstract: "Toronto’s homeless population experienced changes between 2018 and 2024, with key trends in age, gender, and chronic homelessness. Using data from the Toronto Shelter System, this paper examines these trends and the effects of the COVID-19 pandemic. The results show that males aged 35-44 are the largest demographic, and the number of chronically homeless individuals has increased. Understanding these dynamics is essential for developing more effective policies to address homelessness and improve long-term housing stability."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| 
#### Workspace setup ####
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)

@@

In this essay, @sec-data-overview explains the background and necessity of using this dataset, followed by an explanation of the variables mentioned and the methods used. @sec-data-results shows the changes in the number of homeless individuals, along with the proportions and trends over time for each age group, gender, and chronic homelessness. Graphs and related explanations are provided. @sec-discussion further analyses the information provided in the figures and describes the weaknesses of the study.

@@

All processes are performed through the `R` language [@citeR]. Data is simulated, tested, and cleaned using the `tidyverse` [@tidyverse] packages, which also facilitate downloading data through `opendatatoronto` [@citeopendatatorontoR]. Unneeded variables are removed and renamed using `dplyr` [@dplyr] and `tidyr` [@tidyr], while graphs are plotted with `ggplot2` [@ggplot2]. The `knitr` [@knitr] package is used to generate the final PDF report.

@@

@fig-total-population illustrates the fluctuation in Toronto's homeless population from 2018 to 2024. The chart shows cyclical rises and falls, with more pronounced declines around 2020 and 2021, followed by a steady increase. The population reaches its highest point in 2024. Throughout this period, the homeless population varies between approximately 7,500 and 11,000 individuals.

@@

@fig-chronic-population displays the changes in the number and proportion of chronic and non-chronic homeless people in Toronto between 2018 and 2024. The number of chronically homeless has been growing over this time, with over 30,000 people in 2018, and by 2023, the number has gone to about 60,000, nearly doubling. The slight drop in numbers in 2024 is due to the fact that the statistics were last updated on September 17, 2024, and it is projected that by the end of the year, the number will have exceeded the 2023 one. Similarly, the percentage of chronically homeless is growing every year, from 29.7% in 2019 to 63.7% in 2024. Note that in 2019 to 2020 and 2023 to 2024, the rate grows faster, by about 10 percent.
